---
title: "Customizable Forecast Tool"
author: "Jan Kozak"
date: "2024-07-30"
output:
  flexdashboard::flex_dashboard:
    theme:
      bootswatch: flatly
    orientation: rows
    source_code: embed
runtime: shiny
---

``` {r, echo = FALSE, warning=FALSE, message=FALSE}
library(flexdashboard)
library(scales)
library(shinyjs)
library(tidyverse)
library(forecast)
library(plotly)

useShinyjs()

# Load sales data
fsales <- "https://raw.githubusercontent.com/multidis/hult-inter-bus-reports-r/main/forecasting/sales_weekly.csv"
sales <- read_csv(fsales)

# Get the latest week
nweek_now <- max(sales$Week)
```

``` {r echo = FALSE}
tags$style(HTML("
  .hidden {
    display: none;
  }
  .statistics {
    margin-top: 20px;
    padding-top: 20px;
    border-top: 1px solid #ccc;
  }
  .custom-header {
    background-color: #2c3e50;
    color: white;
    padding: 10px;
    text-align: center;
  }
  .custom-header h1 {
    margin: 0;
  }
  .custom-sidebar {
    background-color: #ecf0f1;
    padding: 20px;
  }
  .custom-sidebar h4 {
    margin-top: 0;
  }
  .statistics-output {
    background-color: #dff0d8;
    padding: 10px;
    border-radius: 5px;
    font-size: 1.2em;
  }
  .statistics-output.previous-year {
    background-color: #2ecc71;
    color: white;
  }
  .statistics-output.next-year {
    background-color: #e74c3c;
    color: white;
  }
  .statistics-output.washed-yellow {
    background-color: #fff9c4;
  }
"))
```


Next Year Forecast {data-icon="fa-chart-line"}
==================================================

This is an adjustable interactive chart depicting the forecasted sales of different stores for various time periods. Feel free to adjust the parameters to fit your needs and analysis. 


## Selectors


``` {r echo = FALSE}
fluidRow(
  column(12,
         div(class = "custom-sidebar",
             selectInput("forecast_type", label = h4("Select Forecast Type"), choices = c("Weekly Forecast" = "weekly", "Quarterly Forecast" = "quarterly"), selected = "weekly"),

             conditionalPanel(
               condition = "input.forecast_type == 'weekly'",
               selectInput("storenum_weekly", label = h4("Select Store"), choices = c("All", 1:max(sales$Store)), selected = 1),
               sliderInput("horizon_weekly", label = h4("Forecast Horizon (weeks)"), min = 2, max = 52, value = 8, step = 1)
             ),

             conditionalPanel(
               condition = "input.forecast_type == 'quarterly'",
               selectInput("storenum_quarterly", label = h4("Select Store"), choices = c("All", 1:max(sales$Store)), selected = 1),
               radioButtons("quarter", label = h4("Select Quarter"), choices = c("Q1", "Q2", "Q3", "Q4"), inline = TRUE)
             ),

             tags$div(class = "hidden", selectInput("storenum", label = "Select store: ", choices = c("All", 1:max(sales$Store)), selected = 1)),

             checkboxInput("show_ci", label = h4("Show Confidence Intervals"), value = FALSE),
             conditionalPanel(
               condition = "input.show_ci == true",
               sliderInput("confidence_level", label = h4("Confidence Level (%)"), min = 50, max = 99, value = 95, step = 5)
             ),
             
             checkboxInput("show_trend", label = h4("Show Trendline"), value = FALSE),

             div(class = "statistics",
                 h4("Total Forecasted Sales:"),
                 div(class = "statistics-output previous-year", textOutput("total_sales")),

                 h4("Mean Forecasted Sales:"),
                 div(class = "statistics-output previous-year", textOutput("mean_sales_forecast")),

                 h4("Median Forecasted Sales:"),
                 div(class = "statistics-output previous-year", textOutput("median_sales_forecast"))
             )
         )
  )
)
```



``` {r echo = FALSE}
output$plotly <- renderPlotly({
  forecast_type <- input$forecast_type
  show_ci <- input$show_ci
  show_trend <- input$show_trend
  conf_level <- input$confidence_level / 100

  if (forecast_type == "weekly") {
    # Weekly forecast logic
    storenum <- input$storenum_weekly
    horiz <- input$horizon_weekly

    sales_data <- if (storenum == "All") {
      sales %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
    } else {
      sales %>% filter(Store == storenum) %>% select(Week, Weekly_Sales)
    }

    sales_ts <- ts(sales_data$Weekly_Sales, frequency = 52)
    arima_model <- auto.arima(sales_ts, seasonal.test = "seas")
    arima_pred <- forecast(arima_model, h = horiz, level = conf_level * 100)

    sales_pred <- data.frame(
      Sales_Predicted = as.numeric(arima_pred$mean),
      Week = 1:horiz,
      Upper = as.numeric(arima_pred$upper),
      Lower = as.numeric(arima_pred$lower)
    )

    p <- sales_pred %>% ggplot(aes(Week, Sales_Predicted)) +
      geom_line(color = "#2c3e50") +
      scale_y_continuous(labels = scales::dollar_format(scale = 1e-6, suffix = "M")) +
      labs(x = "Weeks ahead", y = NULL, title = paste("Sales Forecast for the Next", horiz, "Weeks")) +
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 0, hjust = 0.5, color = "#34495e", size = 14, face = "bold"),
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5, color = "#34495e"),
        axis.title.x = element_text(size = 14, face = "bold", color = "#34495e"),
        axis.title.y = element_text(size = 14, face = "bold")
      )

    if (show_ci) {
      p <- p + geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "blue", alpha = 0.2)
    }
    
    if (show_trend) {
      p <- p + geom_smooth(method = "loess", se = FALSE, color = "red")
    }

  } else if (forecast_type == "quarterly") {
    # Quarterly forecast logic
    storenum <- input$storenum_quarterly
    quarter <- input$quarter

    quarter_weeks <- switch(quarter,
                            "Q1" = 1:13,
                            "Q2" = 14:26,
                            "Q3" = 27:39,
                            "Q4" = 40:52)
    sales_data <- if (storenum == "All") {
      sales %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
    } else {
      sales %>% filter(Store == storenum) %>% select(Week, Weekly_Sales)
    }

    sales_ts <- ts(sales_data$Weekly_Sales, frequency = 52)
    arima_model <- auto.arima(sales_ts, seasonal.test = "seas")
    arima_pred <- forecast(arima_model, h = max(quarter_weeks), level = conf_level * 100)

    sales_pred <- data.frame(
      Sales_Predicted = as.numeric(arima_pred$mean[quarter_weeks]),
      Week = quarter_weeks,
      Upper = as.numeric(arima_pred$upper[quarter_weeks]),
      Lower = as.numeric(arima_pred$lower[quarter_weeks])
    )

    p <- sales_pred %>% ggplot(aes(Week, Sales_Predicted)) +
      geom_line(color = "#2c3e50") +
      scale_y_continuous(labels = scales::dollar_format(scale = 1e-6, suffix = "M")) +
      labs(x = "Weeks ahead", y = NULL, title = paste("Sales Forecast for Quarter", quarter, "of Next Year")) +
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 0, hjust = 0.5, color = "#34495e", size = 14, face = "bold"),
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5, color = "#34495e"),
        axis.title.x = element_text(size = 14, face = "bold", color = "#34495e"),
        axis.title.y = element_text(size = 14, face = "bold")
      )

    if (show_ci) {
      p <- p + geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "blue", alpha = 0.2)
    }
    
    if (show_trend) {
      p <- p + geom_smooth(method = "loess", se = FALSE, color = "red")
    }

  }

  ggplotly(p) %>%
    layout(
      margin = list(b = 80)
    )  # Adjust bottom margin to ensure x-axis labels are visible
})

output$total_sales <- renderText({
  forecast_type <- input$forecast_type
  storenum <- if (forecast_type == "weekly") input$storenum_weekly else input$storenum_quarterly
  horiz <- if (forecast_type == "weekly") input$horizon_weekly else switch(input$quarter, "Q1" = 13, "Q2" = 26, "Q3" = 39, "Q4" = 52)

  sales_data <- if (storenum == "All") {
    sales %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
  } else {
    sales %>% filter(Store == storenum) %>% select(Week, Weekly_Sales)
  }

  sales_ts <- ts(sales_data$Weekly_Sales, frequency = 52)
  arima_model <- auto.arima(sales_ts, seasonal.test = "seas")
  arima_pred <- forecast(arima_model, h = horiz)

  if (forecast_type == "weekly") {
    total_sales <- sum(arima_pred$mean[1:horiz])
  } else {
    quarter_weeks <- switch(input$quarter,
                            "Q1" = 1:13,
                            "Q2" = 14:26,
                            "Q3" = 27:39,
                            "Q4" = 40:52)
    total_sales <- sum(arima_pred$mean[quarter_weeks])
  }

  paste(dollar(total_sales))
})

output$mean_sales_forecast <- renderText({
  forecast_type <- input$forecast_type
  storenum <- if (forecast_type == "weekly") input$storenum_weekly else input$storenum_quarterly
  horiz <- if (forecast_type == "weekly") input$horizon_weekly else switch(input$quarter, "Q1" = 13, "Q2" = 26, "Q3" = 39, "Q4" = 52)

  sales_data <- if (storenum == "All") {
    sales %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
  } else {
    sales %>% filter(Store == storenum) %>% select(Week, Weekly_Sales)
  }

  sales_ts <- ts(sales_data$Weekly_Sales, frequency = 52)
  arima_model <- auto.arima(sales_ts, seasonal.test = "seas")
  arima_pred <- forecast(arima_model, h = horiz)

  if (forecast_type == "weekly") {
    mean_sales <- mean(arima_pred$mean[1:horiz])
  } else {
    quarter_weeks <- switch(input$quarter,
                            "Q1" = 1:13,
                            "Q2" = 14:26,
                            "Q3" = 27:39,
                            "Q4" = 40:52)
    mean_sales <- mean(arima_pred$mean[quarter_weeks])
  }

  paste(dollar(mean_sales))
})

output$median_sales_forecast <- renderText({
  forecast_type <- input$forecast_type
  storenum <- if (forecast_type == "weekly") input$storenum_weekly else input$storenum_quarterly
  horiz <- if (forecast_type == "weekly") input$horizon_weekly else switch(input$quarter, "Q1" = 13, "Q2" = 26, "Q3" = 39, "Q4" = 52)

  sales_data <- if (storenum == "All") {
    sales %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
  } else {
    sales %>% filter(Store == storenum) %>% select(Week, Weekly_Sales)
  }

  sales_ts <- ts(sales_data$Weekly_Sales, frequency = 52)
  arima_model <- auto.arima(sales_ts, seasonal.test = "seas")
  arima_pred <- forecast(arima_model, h = horiz)

  if (forecast_type == "weekly") {
    median_sales <- median(arima_pred$mean[1:horiz])
  } else {
    quarter_weeks <- switch(input$quarter,
                            "Q1" = 1:13,
                            "Q2" = 14:26,
                            "Q3" = 27:39,
                            "Q4" = 40:52)
    median_sales <- median(arima_pred$mean[quarter_weeks])
  }

  paste(dollar(median_sales))
})
``` 

### {.content}

```{r echo = FALSE}
plotlyOutput("plotly", height = "600px", width = "100%")
```

Comparative Analysis {data-icon="fa-balance-scale"}
==================================================
This page compares the sales from previous year with projected sales for the upcoming year. 

## Selectors

```{r echo = FALSE}
fluidRow(
  column(12,
         div(class = "custom-sidebar",
             selectInput("store_comparison", label = h4("Select Store"), choices = c("All", 1:max(sales$Store)), selected = "All"),

             h4("Mean Sales:"),
             div(class = "statistics-output previous-year", textOutput("mean_sales_actual_comparison")),
             div(class = "statistics-output next-year", textOutput("mean_sales_forecast_comparison")),

             h4("Median Sales:"),
             div(class = "statistics-output previous-year", textOutput("median_sales_actual_comparison")),
             div(class = "statistics-output next-year", textOutput("median_sales_forecast_comparison")),

             h4("Highest Growth Quarter:"),
             div(class = "statistics-output washed-yellow", uiOutput("highest_growth_quarter")),

             h4("Lowest Growth Quarter:"),
             div(class = "statistics-output washed-yellow", uiOutput("lowest_growth_quarter"))
         )
  )
)
```

```{r echo = FALSE}
output$mean_sales_actual_comparison <- renderText({
  store <- input$store_comparison
  print(paste("Selected store for mean_sales_actual_comparison: ", store))
  
  actual_sales <- if (store == "All") {
    sales %>% filter(Week <= nweek_now) %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
  } else {
    sales %>% filter(Store == store & Week <= nweek_now) %>% select(Week, Weekly_Sales)
  }
  
  mean_sales <- mean(actual_sales$Weekly_Sales)
  paste("Previous Year: ", dollar(mean_sales))
})

output$mean_sales_forecast_comparison <- renderText({
  store <- input$store_comparison
  print(paste("Selected store for mean_sales_forecast_comparison: ", store))
  
  actual_sales <- if (store == "All") {
    sales %>% filter(Week <= nweek_now) %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
  } else {
    sales %>% filter(Store == store & Week <= nweek_now) %>% select(Week, Weekly_Sales)
  }
  
  actual_ts <- ts(actual_sales$Weekly_Sales, frequency = 52)
  forecast_arima <- forecast(auto.arima(actual_ts), h = 52)
  
  mean_sales_forecast <- mean(forecast_arima$mean)
  paste("Next Year: ", dollar(mean_sales_forecast))
})

output$median_sales_actual_comparison <- renderText({
  store <- input$store_comparison
  print(paste("Selected store for median_sales_actual_comparison: ", store))
  
  actual_sales <- if (store == "All") {
    sales %>% filter(Week <= nweek_now) %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
  } else {
    sales %>% filter(Store == store & Week <= nweek_now) %>% select(Week, Weekly_Sales)
  }
  
  median_sales <- median(actual_sales$Weekly_Sales)
  paste("Previous Year: ", dollar(median_sales))
})

output$median_sales_forecast_comparison <- renderText({
  store <- input$store_comparison
  print(paste("Selected store for median_sales_forecast_comparison: ", store))
  
  actual_sales <- if (store == "All") {
    sales %>% filter(Week <= nweek_now) %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
  } else {
    sales %>% filter(Store == store & Week <= nweek_now) %>% select(Week, Weekly_Sales)
  }
  
  actual_ts <- ts(actual_sales$Weekly_Sales, frequency = 52)
  forecast_arima <- forecast(auto.arima(actual_ts), h = 52)
  
  median_sales_forecast <- median(forecast_arima$mean)
  paste("Next Year: ", dollar(median_sales_forecast))
})
``` 

```{r echo = FALSE}
output$highest_growth_quarter <- renderUI({
  store <- input$store_comparison
  print(paste("Selected store for highest_growth_quarter: ", store))
  
  if (is.null(store) || store == "") {
    return(HTML("No store selected"))
  }

  actual_sales <- if (store == "All") {
    sales %>% filter(Week <= nweek_now) %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
  } else {
    sales %>% filter(Store == store & Week <= nweek_now) %>% select(Week, Weekly_Sales)
  }
  
  if (nrow(actual_sales) == 0) {
    return(HTML("No data available for the selected store"))
  }

  actual_ts <- ts(actual_sales$Weekly_Sales, frequency = 52)
  forecast_arima <- forecast(auto.arima(actual_ts), h = 52)
  
  quarters <- list(
    "Q1" = 1:13,
    "Q2" = 14:26,
    "Q3" = 27:39,
    "Q4" = 40:52
  )
  
  growths <- sapply(names(quarters), function(q) {
    past_sales <- sum(actual_sales$Weekly_Sales[quarters[[q]]], na.rm = TRUE)
    forecast_sales <- sum(forecast_arima$mean[quarters[[q]]], na.rm = TRUE)
    if (past_sales == 0) {
      return(NA)  # Avoid division by zero
    }
    (forecast_sales - past_sales) / past_sales * 100
  })
  
  growths <- na.omit(growths)
  print(paste("Growths for highest_growth_quarter: ", growths))

  if (length(growths) == 0) {
    return(HTML("No growth data available for the selected store"))
  }
  
  max_growth_q <- names(growths)[which.max(growths)]
  max_growth <- growths[max_growth_q]
  
  color <- if (max_growth > 0) "green" else "red"
  HTML(paste0("Highest Growth Quarter: ", max_growth_q, " (", sprintf("<span style='color:%s'>%.2f%%</span>", color, max_growth), ")"))
})

output$lowest_growth_quarter <- renderUI({
  store <- input$store_comparison
  print(paste("Selected store for lowest_growth_quarter: ", store))
  
  if (is.null(store) || store == "") {
    return(HTML("No store selected"))
  }

  actual_sales <- if (store == "All") {
    sales %>% filter(Week <= nweek_now) %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
  } else {
    sales %>% filter(Store == store & Week <= nweek_now) %>% select(Week, Weekly_Sales)
  }
  
  if (nrow(actual_sales) == 0) {
    return(HTML("No data available for the selected store"))
  }

  actual_ts <- ts(actual_sales$Weekly_Sales, frequency = 52)
  forecast_arima <- forecast(auto.arima(actual_ts), h = 52)
  
  quarters <- list(
    "Q1" = 1:13,
    "Q2" = 14:26,
    "Q3" = 27:39,
    "Q4" = 40:52
  )
  growths <- sapply(names(quarters), function(q) {
    past_sales <- sum(actual_sales$Weekly_Sales[quarters[[q]]], na.rm = TRUE)
    forecast_sales <- sum(forecast_arima$mean[quarters[[q]]], na.rm = TRUE)
    if (past_sales == 0) {
      return(NA)
    }
    (forecast_sales - past_sales) / past_sales * 100
  })
  growths <- na.omit(growths)
  print(paste("Growths for lowest_growth_quarter: ", growths))

  if (length(growths) == 0) {
    return(HTML("No growth data available for the selected store"))
  }
  
  min_growth_q <- names(growths)[which.min(growths)]
  min_growth <- growths[min_growth_q]
  
  color <- if (min_growth > 0) "green" else "red"
  HTML(paste0("Lowest Growth Quarter: ", min_growth_q, " (", sprintf("<span style='color:%s'>%.2f%%</span>", color, min_growth), ")"))
})
```

``` {r echo = FALSE}
output$comparison_plot <- renderPlotly({
  store <- input$store_comparison
  print(paste("Selected store for comparison_plot: ", store))
  
  actual_sales <- if (store == "All") {
    sales %>% filter(Week <= nweek_now) %>% group_by(Week) %>% summarise(Weekly_Sales = sum(Weekly_Sales))
  } else {
    sales %>% filter(Store == store & Week <= nweek_now) %>% select(Week, Weekly_Sales)
  }
  
  actual_ts <- ts(actual_sales$Weekly_Sales, frequency = 52)
  forecast_arima <- forecast(auto.arima(actual_ts), h = 52)
  
  future_weeks <- seq(nweek_now + 1, nweek_now + 52, by = 1)
  forecast_sales <- data.frame(
    Week = future_weeks,
    Weekly_Sales = as.numeric(forecast_arima$mean)
  )
  
  actual_sales <- actual_sales %>%
    mutate(Calendar_Week = Week %% 52 + 1)
  forecast_sales <- forecast_sales %>%
    mutate(Calendar_Week = Week %% 52 + 1)
  
  comparison_data <- bind_rows(
    actual_sales %>% mutate(Type = "Previous Year"),
    forecast_sales %>% mutate(Type = "Next Year")
  )
  
  p <- ggplot(comparison_data, aes(x = Calendar_Week, y = Weekly_Sales, color = Type)) +
    geom_line(linewidth = 1.2) +
    scale_color_manual(values = c("Previous Year" = "#2ecc71", "Next Year" = "#e74c3c")) +
    scale_y_continuous(labels = scales::dollar_format(scale = 1e-6, suffix = "M")) +
    labs(x = "Weeks of Calendar Year", y = NULL, title = "Previous Year vs Next Year Projected Weekly Sales") +
    theme_minimal() +
    theme(
      plot.title = element_text(color = "#2c3e50", size = 20, face = "bold", hjust = 0.5),
      axis.title.x = element_text(color = "#2c3e50", size = 14, face = "bold"),
      axis.title.y = element_text(size = 14, face = "bold"),
      axis.text = element_text(color = "#34495e"),
      legend.title = element_text(color = "#2c3e50", size = 14, face = "bold"),
      legend.text = element_text(color = "#2c3e50", size = 12)
    )
  
  ggplotly(p) %>%
    layout(legend = list(title = list(text = "Legend")))
})
```

### {.content}

``` {r echo = FALSE}
plotlyOutput("comparison_plot", height = "600px", width = "100%")
``` 
